Linear instability underlying synchronous waving of marine grass in flow

### 2014-09-03 Paper/grass.tex
abstact update september 3rd

The spontaneous waving of marine grass is thought to be due to a Kelvin-Helmholtz instability resulting from an inflection point near the canopy top in the flow profile.
 Our analysis of flow through grass bed predicts a critical threshold condition for instability and associated waving frequency
which compares well with experimental observations. We show that in the limit of a dense grass bed, two distinct unstable modes (Mode-1, Mode-2) can be asymptotically distinguished.
while Mode-1 shares many characteristics with the Kelvin-Helmholtz instability where shear layer near canopy top plays a critical role, 
Mode-2 can be thought of as instability of inviscid flow with dense grass bed for which shear 
layer near grass tip doesn't play a critical role. We further show that in canopy with high submergence ratio
Kelvin-Helmholtz insatbility is damped due to grass drag. In the parameter regime typical of marine grass beds, the two modes merge and are indistinguishable.

@@

\end{abstract}
\maketitle

%\section{Introduction}
Sea grasses exhibit rich set of dynamics due to their interaction with the flow of water and can considerably affect surrounding hydrodynamic conditions.
Resulting changes in hydrodynamic conditions can influence number of environmental variables and processes such as transport of sediments, contaminants, dissolved oxygen, plant growth, and biomass production, etc \cite{Fonseca87,Nepf99,Nepf2012}. 
Flexible grass canopies respond to steady currents in the form of coherent large amplitude oscillations, known as mo-nami \cite{AckermanOkubo93}. Such motion is considered to play a central role in the recruitment of microscopic marine organisms, such as blue mussel larvae\cite{Grizzle96}.
The hydrodynamic mechanism underlying monami is the focus of this paper.

>>>>

%The spontaneous waving of marine grass is thought to be due to a Kelvin-Helmholtz instability resulting from an inflection point near the canopy top in the flow profile.
We study the mechanism behind synchronous waving of marine grass, known as Mo-nami. We do linear stability analysis of a mathematical model which considers grass to exert a continuum drag on the flow.
Our analysis of flow through grass bed predicts a critical threshold condition for instability and associated waving frequency
which compares well with experimental observations. Our analysis further predicts two distinct mode of instability which might lead to Mo-nami.
One of these mode shares many characteristics with the Kelvin-Helmholtz instability, we termed this mode as ``Mode 1''. The other mode can be thought of as instability of 
inviscid flow above grass coupled with the porous flow through grass bed, we refer to this mode as ``Mode 2''.  In the limit of dense grass bed these two modes can be asymptotically 
distinguished. In the parameter regime typical of marine grass beds, the two modes merge and are indistinguishable.

@@

\end{abstract}
\maketitle
%\section{Introduction}
Sea grasses exhibit rich set of dynamics due to their interaction with the flow of water and can considerably affect surrounding hydrodynamic conditions.
Resulting changes in hydrodynamic conditions can influence number of environmental variables and processes such as transport of sediments, contaminants, dissolved oxygen, plant growth, and biomass production, etc \cite{Fonseca87,Nepf99,Nepf2012}. 
Flexible grass canopies respond to steady currents in the form of coherent large amplitude oscillations, known as mo-nami \cite{AckermanOkubo93}. Such motion is considered to play a central role in the recruitment of microscopic marine organisms, such as blue mussel larvae\cite{Grizzle96}.
The hydrodynamic mechanism underlying monami is the focus of this paper.
